\chapter{Introduction}\label{chap:introduction}

\section{Background}

The Republic of Moldva, a small country in Eastern Europe that borders Romania and Ukraine, with a population of 2.4 million people, has faced significant economic and political challenges since gaining independence in 1991 \parencite{mdpop, mdpoverty}. Despite its struggles, Moldova has made substantial progress in its digital transformation initatives aimed at modernising public services and improving citizens' quality of life \parencite{mdega}. A notable example is the Citizen's Government Portal (MCabinet), which provides digital access to all citizens via electronic signature to government-held information and select services \parencite{mdcabinet}. Continuing in this direction, Moldova's Cabinet of Ministers has recently approved the `Digital Transformation Strategy of the Republic of Moldova for 2023--2030', which aims to transform the country into a digital society by 2030, with the ultimate goal of having `all public services available in a digitalized format' \parencite{mdstrategy}.

\section{Problem Statement}

In parallel to government service transformations, the healthcare sector has also seen some digitalisation efforts, evidenced by the introduction of an electronic health record system in 15 hospitals across the country in 2017, called `Sistemul informațional automatizat „Asistența Medicală Spitalicească” (SIA AMS)' \parencite{mdehr}. While the system has improved practitioners' access to patient information, the system hasn't been updated since its inception in 2017 and several challenges persist in 2024.

The system's user experience represents its main limitation, with an outdated interface that is difficult to navigate and fails to meet modern accessibility standards. Additionally, it is only accessible via Internet Explorer or a legacy version of Microsoft Edge, with no support for other browsers or devices \parencite{mdehr}.

Another fundamental challenge is the siloing of electronic health records systems across the country. The lack of a nationally-wide integrated system, such as the NHS in the UK, has resulted in fragmented patient data spread across individual healthcare institutions, with no centralised access available \parencite{mdehr}. In turn, this fragmentation has created an overreliance on paper-based or manual processes, which are time-consuming and prone to human error, hindering the quality of care provided to patients.

Moldova's precarious economic situation further complicates any efforts to modernise the current healthcare system. As such, Moldovan citizens are incresingly turning to private digital initiatives and local startups that offer innovative solutions, as evidenced by recent developments showcased at events like the Digital Health Forum 2024 \parencite{mdstartup}.

\section{The Client}

The department of external relations of `Nicolae Testemiteanu' State University of Medicine and Pharmacy in Moldova (USMF) will serve as this project's client. USMF is a public institution that offers a range of medical programs, including medicine, dentistry and pharmacy \parencite{mduni}. Many of the faculty at USMF are also practicing doctors at hospitals and clinics across Moldova, and have first-hand experience with the current digital systems used in both public and private medical institutions.

Members of the department of external relations have expressed their interest to investigate potential technology-enabled solutions to improve the healthcare sector in Moldova. They have requested for a prototype of a digital system that will either replace or complement existing technological healthcare offerings in Moldova. 

\section{Project Objectives}\label{sec:objectives}

This project aims to address some the challenges faced by the healthcare sector in Moldova, focusing on providing a working, high-fidelity prototype. This will be achieved by a combination of market research and stakeholder interviews, followed by a literature review of existing technologies that will power the prototype, and finally, the design and development of a working prototype. The project's ultimate goal is to create a viable solution that will establish a foundation for potential future broader implementation within Moldova's healthcare sector.


As such, the objectives of the project are as follows:
\begin{enumerate}   
    \item Identify 2 to 4 stakeholders across diverse healthcare perspectives to provide insights into current systems used in Moldova's healthcare sector.
    \item Conduct stakeholder interviews to gather information about current systems used, their challenges and requirements.
    \item Carry out a literature review to research the appropriate technologies and project management methodologies for developing a modern EHR system.
    \item Explore at least 2 existing EHR systems and identify their strengths and weaknesses.
    \item Design and develop a functional EHR system prototype based on the requirements gathered.
    \item Ensure the prototype can be used as a proof of concept to secure funding and support for further development.
\end{enumerate}